Scalable Clustering of Genotype Information using MapReduce

### 2015-03-06 doc/publications/journal2014/paper.tex
updated sections

\newcommand{\OnePhaseone}{phase1\_chr1}
\newcommand{\NinteenPhaseone}{phase1\_chr19}
\newcommand{\SevenPhaseone}{phase1\_chr1-7}
\newcommand{\FullPhaseone}{phase1\_chr1-22}
\newcommand{\OnePhasethree}{phase3\_chr1}
\newcommand{\ThreePhasethree}{phase3\_chr1-3}
\newcommand{\FullPhasethree}{phase3\_chr1-22}

@@

% COMPARISON TO R
\subsection*{Clustering in Mahout requires less memory than in R}
In this section, we compare the runtime and resource requirements between the k-means clustering implementation in R and Mahout for the small \NinteenPhaseone{} dataset (see Methods). 
We measured the runtime separately for pre-processing stage as well as the clustering stage. 
%DB: can you explain 'container' or refer to methods where you have more room for discussing
%DB: is this addressed?
With Hadoop, pre-processing the data takes approximately 6 minutes, with an allocation of up to 3GB memory per 'container'. The resulting data is stored as a Hadoop 'SequenceFile' of about 835MB in size.
Conversely,  pre-processing the data to R's in-memory matrix takes 56 minutes and required 17.7GB memory. 
K-means clustering the data, however, is faster with R, using 7s, whereas Mahout takes just under 9 minutes. See Table~\ref{Tab:02}. \\
Due to the relatively small number of variants, cluster performance is relatively poor for both methods, with the best Adjusted Rand Index score for clustering individuals in one of 4 populations (see Methods) being 0.705 in Mahout and 0.614 in R (see Table~\ref{Tab:02}).

Although the overall Mahout pipeline is about 40 minutes faster than R's, the actual k-means clustering takes two orders of magnitude longer in Mahout. 
This is due to the additional overhead involved in spawning the map and reduce processes, and distributing the data between nodes.
R, on the other hand, already has the entire matrix in-memory, on a single machine. 
Since no disk-access or networking is required the clustering can be performed faster. However, the amount of memory available on a single machine will quickly become a limiting factor. 
As an example, the entire set of 1000 Genome Project phase3 variants could require approximately 1.6TB to be completely loaded into memory as a double-precision matrix (see Table~\ref{Tab:01}). This is based on the equation:
$\frac{n \times m \times 8\text{bytes}}{1\text{GB}}$, where $n \times m$ are the matrix dimensions, 8bytes is the memory a double occupies (i.e. each matrix element), and  1GB is $10^9$bytes.

% DB: not clear you need to spell out your argument a bit more...
% AO: Fixed maybe
% DB: good explanation but I'm not sure what it does here (maybe reuse above or in methods)
% DB: Delete this ?
%The primary benefit of Hadoop is its scalability. This is achieved through MapReduce jobs, which allow the task at hand to be separated into 'containers'.
%Although in the previous example, Hadoop had access to 36GB of memory distributed over 3 machines, the size of each container was only 3GB.
%Each of these containers is effectively a self-contained job that is independent from any other container.
%Therefore, the more resources that the user has available, the more containers that can be run in parallel, either on the same machine or separate, and in turn decreasing the run-time of the overall job.
%This means that our Hadoop job can scale from a single machine with minimal resources (albeit with a longer run-time), to a huge compute cluster with terabytes of RAM.
%This benefit allows Hadoop jobs to not only scale vertically, i.e. to a larger single machine as R can, but also horizontally, to a cluster of many machines.

% (1) SCALING UP FEATURES IN MAHOUT
\subsection*{Hadoop scales to genome-wide clustering}
In this section we demonstrate Mahout's ability to pre-process and cluster genome-wide variants and its effect on clustering quality.
We focus on the larger \FullPhasethree{} dataset in this section which has 79 million variants.
However, we begin with a subset of this dataset, \OnePhasethree{}. Although our algorithm completes successfully, it achieves a relatively low adjusted Rand index of 0.622.

Next we increase the input to include the first three chromosomes (\ThreePhasethree{}), which has just under 20 million variants. This completes successfully and results in an adjusted Rand index of 0.825.
Finally we include all 22 autosomes (\FullPhasethree{}) as input to our algorithm. Although the adjusted rand index isn't much higher than the previous example, at 0.84, the pre-processing and k-means complete successfully.
-Conclusionary statement-

% This section feels like a listing numbers rather than a "story". Maybe having a plot we can make a more general statement. So can you make a plot #features vs memory and in a separate on vs time with different colours showing pre-post processing. Or one plot memory vs time with colours as features and shape as pre-post ?

%% OLD
%%We also ran jobs using intermediately sized phase1 datasets (see Table~\ref{Tab:02}) to demonstrate the different container sizes required by pre-processing and k-means clustering, and time required.
%On the \OnePhaseone{} dataset, pre-processing takes just under 10 minutes and clustering takes approximately 12 minutes with the same container size as with the \NinteenPhaseone{} dataset.
%We then increased the input to include the first seven chromosomes (\SevenPhaseone{}). In this case, pre-processing time increased to approximately 45 minutes, and clustering to 3 hours and 28 minutes.
%Our pre-processing algorithm was successful with the same container size as the previous examples. However, we increased both the map and reduce container size for the clustering step to 4GB to accommodate the increased vector size.
%
%Finally we used the entire phase one dataset as input. Although pre-processing now takes 85 minutes, it was again successful with the same sized containers as previously.
%Clustering, however, not only required larger containers, but also took over 7 hours to complete.
%
%This substantial time increase is in part due to the larger container size. With each container allocated 6GB of memory, our cluster can only run up to six containers simultaneously, as opposed to 24 1.5GB containers (from the smaller examples).
%This restriction in simultaneous containers is due to the limited memory available on our cluster.
%
%% mentioning phase 3 here makes it difficult to understand.
%Although we compare phase3 datasets in the next section because they have a larger number of samples, the \FullPhasethree{} dataset also has the largest feature size, so we mention that in this section to demonstrate Hadoops ability to scale with feature-size.
%The most notable difference is the time required to complete clustering, at just over 21 hours. Although this is a substantial increase in time, the job still completes successfully. As we mention in a later section, this time can be reduced dramatically on a larger cluster.
%
%%DB how else would you expect memory to grow 
%This section demonstrates Hadoop's ability to cope with genome wide information, which predominantly results in a runtime increase. The memory requirements are also increased, however the memory required increases in proportion to vector length,
%rather than the entire size of the dataset. This means we can successfully run a job on the variants from an entire genome on our cluster, or potentially an even smaller cluster, as long as 6GB is available for containers.
%rather than memory increase which quickly become the limiting factor as demonstrated by R's inability to process the data. 
% /OLD

% (2) SCALING UP SAMPLES IN MAHOUT
\subsection*{Hadoop scales effortlessly to more than twice the number of samples}

In this section we investigate Hadoop's ability to scale with more samples. Mahout's k-means algorithm performs by splitting clustering jobs up into separate, smaller tasks, where each task calculates the distance between a subset of samples and the k-means centroids.
Hadoop creates a 'container' for each task, and processes them iteratively until every task is complete.
We discuss containers in greater detail in the next section, however a container is essentially a pre-defined allocation of memory and CPU cores.
In the previous section we increased the sample size. Larger samples have an increased dimensionality, and hence require more memory per container, and also more time to complete.
In this case, however, we increase the number of samples. With more samples, the k-means algorithm will divide the job up into more tasks, however the memory requirement of each task, and hence, container, should remain the same because the dimensionality is the same.
Instead the job will take more time because there are more tasks for Mahout to complete.
To demonstrate this, we compare k-means clustering of the 2504 individuals from the phase3 dataset compared to the 1092 individuals in the phase1 dataset.
We specifically use the \SevenPhaseone{} and \ThreePhasethree{} datasets because of their similar feature size.
The algorithm completed successfully with the same sized containers for both datasets, even with more than twice the number of samples in the larger dataset.
However, the \ThreePhasethree{} dataset required about twenty minutes more time than the smaller \SevenPhaseone{} dataset.
This demonstrates Hadoop's strength of scaling with sample size. Rather than the larger comparison requiring us to allocate more resources allocated to containers, it simply required more time. Therefore, increasing the sample size of a job should not require reconfiguration of Hadoop or a larger cluster.

@@

% (3) CLUSTER AND CONTAINER SIZE
\subsection*{Cluster size and containers}
When designing a MapReduce job, it is important that the user specify container size correctly. Overallocation of resources can result in wasted resources, whereas under allocation can result in failed jobs.
To elaborate, a container is essentially an abstraction of resources, i.e. memory and CPU cores. After Hadoop divides a job up into tasks, each task runs in parallel in a container. In addition, each container is independent of other containers.
The resources Hadoop allocates to each container is a predefined value, specified in its configuration. According to the Cloudera documentation, the general rule of thumb is that one CPU core is allocated to each container, and the available memory per node should be divided equally between each of those containers. For example, on a node with 8 CPU cores and 32GB memory, each container should be allocated 1 CPU core and 4GB memory. Effectively allowing a maximum of eight 4GB containers to run in parallel on that node.

@@

To demonstrate the importance of specifying these parameters correctly, we ran the identical jobs, using different Hadoop configurations. The \OnePhaseone{} dataset only requires 1.5GB memory per container, as we demonstrated previously.
Increasing the map container size to 4GB results in the job taking an additional 5 minutes (see Table \ref{Tab:03}).
This larger container size is still too small for the \FullPhasethree{} dataset, which we demonstrated to require 6GB memory per container. Running a job on this dataset with 4GB memory per container results in the job failing.

One of the main selling points of Hadoop, is that it will run on commodity hardware. For this reason we used a relatively small cluster for our benchmarks. Although our cluster uses VMs on Microsoft Azure, it would be relatively simple for a lab to
create a similar Hadoop setup in-house. However, to demonstrate the scaling ability of Hadoop and our MapReduce algorithms, we also ran the clustering jobs on a much larger development Hadoop cluster. As presented in Table \ref{Tab:03}, the \OnePhaseone{} job saw an improvement of approximately 70\%.
The \FullPhasethree{} example, on the other hand, was 78\% faster.
Although both jobs were more than twice as fast on the larger cluster, the speed doesn't scale relative to the size of the cluster. This is due to a limitation in these examples, where the number of reduces in k-means clustering is limited to the number of k-means clusters (k).
A task where k is five, for example, will only be able to spawn five reduce containers, no matter how many resources are available. Although sometimes inevitable, depending on the task, this bottleneck of maximum simultaneous reduce tasks is an important consideration to take into account when designing MapReduce jobs.

@@

\subsection*{Clustering quality}
We scored the clusters with the Adjusted Rand index~\cite{Hubert1985}. 
This index compares two different cluster sets, here the prediction $X = \{ X_1, X_2, \ldots , X_r \}$ and the annotated data $Y = \{ Y_1, Y_2, \ldots , Y_s \}$, and assigns a value based the overlap between $X$ and $Y$ as captured in a contingency table $\left[n_{ij}\right]$ where each entry $n_{ij}$ denotes the number of objects in common between $X_i$ and $Y_j : n_{ij}=|X_i \cap Y_j|$. 
The Adjusted Rand index is then 
{\tiny
\begin{eqnarray*}
ARI=\frac{\sum_{ij}{{n_{ij}\choose 2}} - \left[ \sum_{i}{{a_i\choose2}} \sum_{j}{{b_i\choose2}} \right] / {n\choose2}}{\frac{1}{2} \left[ \sum_{i}{{a_{i}\choose 2} + \sum_{j}{{b_{j}\choose 2}}} \right] - \left[ \sum_{i}{{a_{i}\choose 2} + \sum_{j}{{b_{j}\choose 2}}} \right] / {n\choose2}} 
\end{eqnarray*}
}
where $a$ and $b$ are the sums over the rows and columns of the contingency table.

@@

\begin{table*}[!t]
\processtable{Resources allocated and time taken.\label{Tab:02}}
{\begin{tabular}{ll|cc|cc|rr|c}\toprule
Tool & Dataset & \multicolumn{2}{ c |}{Preprocesing (GB)} & \multicolumn{2}{ c |}{Clustering (GB)} & \multicolumn{2}{ c |}{Time} & Accuracy\\
& & map & reduce  & map & reduce  & pre-processing & clustering & \\\midrule
        R &\NinteenPhaseone{} & - & - & - & - & 55m 41s & 7s &0.614\\
        Mahout & \NinteenPhaseone{} & 1536 & 3072 & 1536 & 3072 & 6m 10s & 8m 32s & 0.705\\
        & \OnePhaseone{} & 1536 & 3072 & 1536 & 3072 & 9m 52s & 12m 49s & 0.798\\
        & \SevenPhaseone{} & 1536 & 3072 & 4096 & 4096  & 43m 56s & 3h 28m & 0.813\\
       % & \FullPhaseone{} & 1536 & 3072 & 6144 & 6144  & 85m 29s & 7h 43m & 0.723\\
        & \OnePhasethree{} & 1536 & 3072 & 1536 & 3072 & 23m 27s & 21m 14s & 0.622\\
        & \ThreePhasethree{} & 1536 & 3072 & 4096 & 4096 & 63m 49s & 4h 21m & 0.825\\
        & \FullPhasethree{} & 1536 & 3072 & 6144 & 6144 & 5h 36m & 21h 5m &0.840\\\botrule
\end{tabular}}{The resources allocated to each MapReduce yarn-container and time taken for the job(s) to complete.}
\end{table*}

\begin{table*}[!t]
\processtable{Resources allocated and time taken.\label{Tab:03}}
{\begin{tabular}{l|cc|cc|r}\toprule
Dataset & \multicolumn{2}{ c |}{Available Resources} & \multicolumn{2}{ c |}{Container Size (GB)} & Time\\
& CPU cores & Memory (GB)  & map & reduce & \\\midrule
        \OnePhaseone{} & 24 & 36 & 1536 & 3072 & 12m 49s\\
        \OnePhaseone{} & 24 & 36 & 4096 & 4096  & 17m 47s\\
        \OnePhaseone{} & 448 & 1310 & 1536 & 4096  & 5m 22s\\
        \FullPhasethree{} & 24 & 36 & 6144 & 6144 & 21h 4m\\
        \FullPhasethree{} & 448 & 1310 & 6144 & 6144 & 4h 42m\\\botrule
\end{tabular}}{The resources allocated to each MapReduce yarn-container and time taken for the job(s) to complete.}
\end{table*}

>>>>

\newcommand{\OnePhaseone}{phase1\_chr1}
\newcommand{\NinteenPhaseone}{phase1\_chr19}
\newcommand{\SevenPhaseone}{phase1\_chr1-7}
\newcommand{\FullPhaseone}{phase1\_chr1-22}
\newcommand{\OnePhasethree}{phase3\_chr1}
\newcommand{\ThreePhasethree}{phase3\_chr1-3}
\newcommand{\FullPhasethree}{phase3\_chr1-22}
\newcommand{\ARI}{Adjusted Rand Index}

@@

% COMPARISON TO R
\subsection*{Clustering in Mahout requires less memory than in R}
In this section, we compare the runtime and resource requirements between the k-means clustering implementation in R and Mahout for the small \NinteenPhaseone{} dataset (see Methods). 
We measured the runtime separately for pre-processing stage as well as the clustering stage.

%With Hadoop, pre-processing the data takes approximately 6 minutes, with an allocation of up to 3GB memory per 'container'. The resulting data is stored as a Hadoop 'SequenceFile' of about 835MB in size.
With Hadoop, pre-processing the data takes approximately 6 minutes, with an allocation of up to 3GB memory. 
The resulting data is stored as a Hadoop 'SequenceFile' of about 835MB in size.
Conversely,  pre-processing the data to R's in-memory matrix takes 56 minutes and required 17.7GB memory. 
K-means clustering the data, however, is faster with R, using 7s, whereas Mahout takes just under 9 minutes (see Table~\ref{Tab:02}). 
This is due to the additional overhead involved in spawning the map and reduce processes, and distributing the data between nodes.
R, on the other hand, already has the entire matrix in-memory, on a single machine and since no disk-access or networking is required the clustering can be performed faster. 
Due to the relatively small number of variants, cluster accuracy is relatively poor for both methods, with the best \ARI{} score for clustering individuals in one of 4 populations (see Methods) being 0.705 in Mahout and 0.614 in R (see Table~\ref{Tab:02}).

%Although the overall Mahout pipeline is about 40 minutes faster than R's, the actual k-means clustering takes two orders of magnitude longer in Mahout. 
%This is due to the additional overhead involved in spawning the map and reduce processes, and distributing the data between nodes.
%R, on the other hand, already has the entire matrix in-memory, on a single machine. 
%Since no disk-access or networking is required the clustering can be performed faster.

While the overall processing time is already 40 minutes faster in Mahout than R, the real bottleneck for larger dataset will be the amount of memory available on a single machine. 
As an example, the entire set of 1000 Genome Project phase3 variants could require approximately 1.6TB to be completely loaded into memory as a double-precision matrix (see Table~\ref{Tab:01}). This is based on the equation:
$\frac{n \times m \times 8\text{bytes}}{1\text{GB}}$, where $n \times m$ are the matrix dimensions, 8bytes is the memory a double occupies (i.e. each matrix element), and  1GB is $10^9$bytes.

Therefore, the data processed in the subsequent sections are too large to be processed with standard R, we hence we focus on evaluating different settings of the Mahout implementation only.

% (1) SCALING UP FEATURES IN MAHOUT
\subsection*{Hadoop scales to genome-wide clustering}
%In this section we demonstrate Mahout's ability to pre-process and cluster genome-wide variants and its effect on clustering quality.
%We focus on the larger \FullPhasethree{} dataset in this section which has 79 million variants.
%However, we begin with a subset of this dataset, \OnePhasethree{}. Although our algorithm completes successfully, it achieves a relatively low adjusted Rand index of 0.622.
%Next we increase the input to include the first three chromosomes (\ThreePhasethree{}), which has just under 20 million variants. This completes successfully and results in an adjusted Rand index of 0.825.
%Finally we include all 22 autosomes (\FullPhasethree{}) as input to our algorithm. Although the adjusted rand index isn't much higher than the previous example, at 0.84, the pre-processing and k-means complete successfully.
%-Conclusionary statement-

In this section we demonstrate Mahout's ability to cluster genome-wide variants. 
To demonstrate accuracy-gain and resource-consumption we evaluate whole-genome data (\FullPhasethree{}), which contains 79 million variants, as well as two subsets with 19 (\ThreePhasethree{}) and 6 million (\OnePhasethree{}) variants, respectively.  
As expected, providing variants from only chromosome one (\OnePhasethree{}) achieves a relatively low \ARI{} of 0.622 (see Table~\ref{Tab:02})
Note, phase3 data has one more population group, which explains the decrease in accuracy compared to the previous section.

Increasing the input to include the first three chromosomes (\ThreePhasethree{}) improves the accuracy with an \ARI{} of 0.825 (Table~\ref{Tab:02}).
We are reaching a saturation point for k-means clustering here, as including all 22 autosomes (\FullPhasethree{}) improves the \ARI{} only marginally to 0.84 (Table~\ref{Tab:02}).
%DB is that really GB or should that be 6.1GB?
However, this demonstrates that performing clustering using whole-genome variants on research infrastructure is feasible using Mahout as the resource consumption lies within realistic boundaries. The time consumption increases to 25 hours and memory to 6144 GB (Table~\ref{Tab:02}).

% (2) SCALING UP SAMPLES IN MAHOUT
\subsection*{Hadoop scales effortlessly to more than twice the number of samples}
%In this section we investigate Hadoop's ability to scale with more samples. 
%Mahout's k-means algorithm performs by splitting clustering jobs up into separate, smaller tasks, where each task calculates the distance between a subset of samples and the k-means centroids.
%Hadoop creates a 'container' for each task, and processes them iteratively until every task is complete.
%In the previous section we increased the sample size. 
%Larger samples have an increased dimensionality, and hence require more memory per container, and also more time to complete.
%In this case, however, we increase the number of samples. 
%With more samples, the k-means algorithm will divide the job up into more tasks, however the memory requirement of each task, and hence, container, should remain the same because the dimensionality is the same.
%Instead the job will take more time because there are more tasks for Mahout to complete.
%In the previous section we increased the sample size. 
%Larger samples have an increased dimensionality, and hence require more memory per container, and also more time to complete.
%In this case, however, we increase the number of samples. 
%With more samples, the k-means algorithm will divide the job up into more tasks, however the memory requirement of each task, and hence, container, should remain the same because the dimensionality is the same.
%Instead the job will take more time because there are more tasks for Mahout to complete.

Hadoop's main advantage over other parallelisation techniques is its ability to scale to more samples without altering the code. 
Mahout's k-means algorithm is hence structured such that it splits the clustering problem up into separate, smaller tasks, where each task calculates the distance between a subset of samples and the k-means centroids. 
The automatic scaling is then achieved by Hadoop creating 'containers' for each task, and processes them iteratively until every task is complete.
We discuss containers in greater detail in the next section, however, a container is essentially a pre-defined allocation of memory and CPU cores.
By increasing the number of samples, the k-means algorithm will simply divide the job up into more tasks, which enables the memory requirement of each task to stay the same. 
This is because the dimensionality is constant, compared to the previous section where we increased the number of variants and observe a subsequent increase in memory.

To demonstrate this, we compare k-means clustering of the 2504 individuals from the phase3 dataset compared to the 1092 individuals in the phase1 dataset.
We specifically use the \SevenPhaseone{} and \ThreePhasethree{} datasets because of their similar feature size.
%The algorithm completed successfully with the same sized containers for both datasets, even with more than twice the number of samples in the larger dataset.
%However, the \ThreePhasethree{} dataset required about twenty minutes more time than the smaller \SevenPhaseone{} dataset. 
The algorithm completed successfully with the same sized containers for both datasets, however, the \ThreePhasethree{} dataset required about twenty minutes more time than the smaller \SevenPhaseone{} dataset (Table~\ref{Tab:02}). 
This demonstrates Hadoop's strength of scaling with sample size.

%Rather than the larger comparison requiring us to allocate more resources allocated to containers, it simply required more time. Therefore, increasing the sample size of a job should not require reconfiguration of Hadoop or a larger cluster.

@@

% (3) CLUSTER AND CONTAINER SIZE
\subsection*{Cluster size and containers}
In the previous section we observed that while the memory remained constant the runtime increased for larger sample sizes. 
On a larger cluster the runtime would also have remained constant. 
%When designing a MapReduce job, it is important that the user specify container size correctly.
However, as research is carried out on resource-finite infrastructure, we discuss in this section how to specify the optimal container size when designing MapReduce jobs to incur only a minimal increase in runtime. 
%Overallocation of resources can result in wasted resources, whereas under allocation can result in failed jobs.

To elaborate, a container is essentially an abstraction of resources, i.e. memory and CPU cores. 
After Hadoop divides a job up into tasks, each task runs in parallel in a container. In addition, each container is independent of other containers.
The resources Hadoop allocates to each container is a predefined value, specified in its configuration. 
% DB reference or version
According to the Cloudera documentation, the general rule of thumb is that one CPU core is allocated to each container, and the available memory per node should be divided equally between each of those containers. 
For example, on a node with 8 CPU cores and 32GB memory, each container should be allocated 1 CPU core and 4GB memory. Effectively allowing a maximum of eight 4GB containers to run in parallel on that node.

@@

To demonstrate the importance of specifying these parameters correctly, we ran the identical jobs, using different Hadoop configurations. 
The \OnePhaseone{} dataset only requires 1.5GB memory per container, as we demonstrated previously.
%Increasing the map container size to 4GB results in the job taking an additional 5 minutes (see Table \ref{Tab:03}).
Increasing the map container size to 4GB results in the job taking an additional 5 minutes (see Table \ref{Tab:03}), because fewer jobs can run in parallel. 
%This larger container size is still too small for the \FullPhasethree{} dataset, which we demonstrated to require 6GB memory per container. 
%Running a job on this dataset with 4GB memory per container results in the job failing.
Attempting to run the larger data set (\FullPhasethree{}) with either 1.5GB or 4GB will result in the job failing, because the feature vector cannot be loaded completely, this requires at least 6GB memory per container and 21h runtime.

One of the main selling points of Hadoop, is that it will run on commodity hardware. 
%For this reason we used a relatively small cluster for our benchmarks. Although our cluster uses VMs on Microsoft Azure, it would be relatively simple for a lab to
%create a similar Hadoop setup in-house. 
To mimic the available resource in a standard resource setting we used only a small VM cluster on Microsoft Azure for the benchmarking in this paper so far.
However, to demonstrate the scaling ability of Hadoop and our MapReduce algorithms, we also ran the clustering jobs on a much larger development Hadoop cluster. 
%As presented in Table \ref{Tab:03}, the \OnePhaseone{} job saw an improvement of approximately 70\%.
%The \FullPhasethree{} example, on the other hand, was 78\% faster.
As presented in Table \ref{Tab:03}, the \OnePhaseone{} job saw an improvement of approximately 70\% when providing 448 cores of 13GB. 
With these resource, the \FullPhasethree{} example, was 78\% faster.

Although both jobs were more than twice as fast on the larger cluster, the speed doesn't scale relative to the size of the cluster. This is due to a limitation in these examples, where the number of reduces in k-means clustering is limited to the number of k-means clusters (k).
A task where k is five, for example, will only be able to spawn five reduce containers, no matter how many resources are available. Although sometimes inevitable, depending on the task, this bottleneck of maximum simultaneous reduce tasks is an important consideration to take into account when designing MapReduce jobs.

@@

\subsection*{Clustering quality}
We scored the clusters with the \ARI{}~\cite{Hubert1985}. 
This index compares two different cluster sets, here the prediction $X = \{ X_1, X_2, \ldots , X_r \}$ and the annotated data $Y = \{ Y_1, Y_2, \ldots , Y_s \}$, and assigns a value based the overlap between $X$ and $Y$ as captured in a contingency table $\left[n_{ij}\right]$ where each entry $n_{ij}$ denotes the number of objects in common between $X_i$ and $Y_j : n_{ij}=|X_i \cap Y_j|$. 
The \ARI{} is then 
{\tiny
\begin{eqnarray*}
ARI=\frac{\sum_{ij}{{n_{ij}\choose 2}} - \left[ \sum_{i}{{a_i\choose2}} \sum_{j}{{b_i\choose2}} \right] / {n\choose2}}{\frac{1}{2} \left[ \sum_{i}{{a_{i}\choose 2} + \sum_{j}{{b_{j}\choose 2}}} \right] - \left[ \sum_{i}{{a_{i}\choose 2} + \sum_{j}{{b_{j}\choose 2}}} \right] / {n\choose2}} 
\end{eqnarray*}
}
where $a$ and $b$ are the sums over the rows and columns of the contingency table.

@@

\begin{table*}[!t]
\processtable{Resources allocated and time taken.\label{Tab:02}}
{\begin{tabular}{ll|cc|cc|rr|c}\toprule
Tool & Dataset & \multicolumn{2}{ c |}{Preprocesing (GB)} & \multicolumn{2}{ c |}{Clustering (GB)} & \multicolumn{2}{ c |}{Time} & Accuracy\\
& & map & reduce  & map & reduce  & pre-processing & clustering & \\\midrule
        R &\NinteenPhaseone{} & - & - & - & - & 55m 41s & 7s &0.614\\
        Mahout & \NinteenPhaseone{} & 1536 & 3072 & 1536 & 3072 & 6m 10s & 8m 32s & 0.705\\
        & \OnePhaseone{} & 1536 & 3072 & 1536 & 3072 & 9m 52s & 12m 49s & 0.798\\
        & \SevenPhaseone{} & 1536 & 3072 & 4096 & 4096  & 43m 56s & 3h 28m & 0.813\\
       % & \FullPhaseone{} & 1536 & 3072 & 6144 & 6144  & 85m 29s & 7h 43m & 0.723\\
        & \OnePhasethree{} & 1536 & 3072 & 1536 & 3072 & 23m 27s & 21m 14s & 0.622\\
        & \ThreePhasethree{} & 1536 & 3072 & 4096 & 4096 & 63m 49s & 4h 21m & 0.825\\
        & \FullPhasethree{} & 1536 & 3072 & 6144 & 6144 & 5h 36m & 21h 5m &0.840\\\botrule
\end{tabular}}{The resources allocated to each MapReduce yarn-container and time taken for the job(s) to complete as well as the accuracy measured as \ARI{}.}
\end{table*}

\begin{table*}[!t]
\processtable{Resources allocated and time taken.\label{Tab:03}}
{\begin{tabular}{cl|cc|cc|r}\toprule
Dataset & \multicolumn{2}{ c |}{Available Resources} & \multicolumn{2}{ c |}{Container Size (GB)} & Time\\
& CPU cores & Memory (GB)  & map & reduce & \\\midrule
        \OnePhaseone{} & 24 & 36 & 1536 & 3072 & 12m 49s\\
        \OnePhaseone{} & 24 & 36 & 4096 & 4096  & 17m 47s\\
        \OnePhaseone{} & 448 & 1310 & 1536 & 4096  & 5m 22s\\
        \FullPhasethree{} & 24 & 36 & 6144 & 6144 & 21h 4m\\
        \FullPhasethree{} & 448 & 1310 & 6144 & 6144 & 4h 42m\\\botrule
\end{tabular}}{The resources allocated to each MapReduce yarn-container and time taken for the job(s) to complete.}
\end{table*}
